FLM for growth as function of SPEI

### 2021-02-12 analysis/paper/vital rate models.Rmd
moved figure 1 and data wrangling out to separate files.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, paged.print=FALSE)
library(tidyverse)
library(here)
library(mgcv)
library(dlnm)
library(bbmle)
library(lubridate)
library(glue)
library(patchwork)
library(sf)
library(tsibble)
library(broom)
library(gratia)
library(latex2exp)

@@

```{r}
#define plotting themes
theme_talk <- theme_bw() + theme(text = element_text(size = 18))
```

@@

```{r data, echo=TRUE}
xa <- read_rds(here("analysis", "data", "derived_data", "xa_spei_lags.rds"))
demog <-
  read_csv(
    here("analysis", "data", "derived_data", "ha_survey.csv"),
    col_names = TRUE,
    cols(plot = col_character(),
         bdffp_reserve_no = col_character())
  )
plot_coords <- read_csv(here("analysis", "data", "raw_data", "plot_coords.csv"))
```

# Join data sets

## Prep precip data

Get only rows for february (the month of the census), and year for joining

```{r}
xa2 <- 
  xa %>%
  filter(month(yearmonth) == 2) %>% 
  mutate(year = year(yearmonth), .after =yearmonth) %>% 
  separate(latlon, into = c("lat", "lon"), sep = "_")
xa2
```

Figure out which sites correspond to which grid cells.  The `xa` lat lon are the centers of grid cells that are 0.25º^2

```{r}
unique(xa2$lat)
mean(c(-2.25, -2.5))
```

So, from -2.25 to -2.5

```{r}
unique(xa2$lon)
mean(c(-59.5, -59.75))
mean(c(-59.75, -60)) # Florestal-CF, 5752, 5751, 5750, 5756, CaboFrio-CF
mean(c(-60, -60.25)) # 2206, 2108, 2107, Dimona-CF, 5753, PortoAlegre-CF
```

```{r}
plot_coords
```

```{r}
demog2 <-
  demog %>% 
  mutate(lon = case_when(
    plot %in% c("Florestal-CF", "5752", "5751", "5750", "5756", "CaboFrio-CF") ~ "-59.875",
    plot %in% c("2206", "2108", "2107", "Dimona-CF", "5753", "PortoAlegre-CF") ~ "-60.125",
    TRUE ~ NA_character_
  ), .before = plot)
demog_full <- 
  left_join(demog2, xa2, by = c("lon", "year")) %>% 
  select(-lon, -lat, -yearmonth)
```

@@

# Inspect raw data

```{r}
demog_plotdf <-  
  demog_post %>% 
  filter(habitat !="10-ha") %>% 
  select(ranch, bdffp_reserve_no, plot, habitat, year, ht, shts, size, surv, flwr) %>% 
  mutate(date = paste(year, "-02-01") %>% ymd(),
         yearmonth = yearmonth(date))
```

## SPEI

```{r}
dates <- range(demog_post$year) %>% paste0(., "-02-01") %>% ymd()
census_dates <- min(demog_post$year)[1]:max(demog_post$year) %>% paste(., "-02-01") %>% ymd()

spei <-
  xa %>%
  select(latlon, yearmonth, spei) %>%
  filter(!is.na(spei)) %>% 
  ggplot(aes(x = yearmonth, y = spei)) + 
  geom_line(aes(group = latlon), alpha = 0.2) +
  stat_summary(geom = "line", fun = "mean") +
  scale_y_continuous("SPEI", expand = expansion(mult = c(0, 0.05), add = 0)) +
  scale_x_yearmonth("Date", limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

spei <- annotate_spei(spei)
spei
```
```{r}
demog_post %>% filter(spei_history[,36] > 2) %>% pull(year) %>% unique()
demog_post %>% filter(year == min(year)) %>% filter(surv == 0)
```

```{r}
demog_post %>% ggplot(aes(x = surv)) + geom_bar() + facet_wrap(~habitat)
```

```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(surv = sum(surv, na.rm = TRUE) / n())
```

Survival nearly identical for 1 ha and CF.

### Timeseries

```{r}
survival <-
  demog_plotdf %>%  
  ggplot(aes(x = yearmonth, y = surv, color = habitat)) +
  stat_summary(geom = "line", fun = "mean", fun.args = list(na.rm = TRUE)) +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(survived)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
survival
# (survival / spei) 
```
Looks like survival went down year after 2003 drought and during 2007 drought.

## Size / growth

```{r}
demog_post %>% ggplot(aes(x = habitat, y = log_size)) + geom_boxplot()
```
```{r}
demog_post %>% 
  group_by(habitat) %>% 
  summarize(size = mean(size, na.rm = TRUE),
            shts = mean(shts, na.rm = TRUE),
            ht = mean(ht, na.rm = TRUE))
```

### Timeseries

```{r}
size <-
  demog_plotdf %>% 
  ggplot(aes(x = yearmonth, y = size, color = habitat)) +
  stat_summary(geom = "line", fun = "mean", fun.args = list(na.rm = TRUE)) +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("Size") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
size
# size/survival/spei
```
Average plant size shrank in 2003 drought, but not 2007 drought

## Flowering

```{r}
demog_post %>% ggplot(aes(x = flwr)) + geom_bar() + facet_wrap(~habitat)
```
### Timeseries

```{r}
flowering <-
  demog_plotdf %>% 
  ggplot(aes(x = yearmonth, y = flwr, color = habitat)) + 
  stat_summary(geom = "line", fun = "mean") +
  stat_summary(geom = "pointrange", fatten = 1) +
  scale_x_yearmonth(limits = dates, breaks = census_dates, date_minor_breaks ="1 month", expand = expansion(mult = 0.02)) +
  scale_y_continuous("P(flowering)") +
  theme_bw() +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), panel.grid.minor.x = element_blank())
flowering
```
```{r}
demog_plotdf %>% filter(year == 2003) %>% group_by(habitat) %>% summarize(flwr = mean(flwr, na.rm = TRUE))
```

```{r}
size/survival/flowering/spei +
  plot_layout(guides = "collect") &
  plot_annotation(tag_levels = "a", tag_suffix = ")") &
  theme(plot.margin = margin(2,1,1,1)) &
  guides(col = guide_legend(title = "Habitat"))
ggsave(here("analysis", "figures", "eda.png"), height = 7, width = 10)
```

@@

combine
```{r}
s_1ha_cbplot + s_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "surv_both_spei.png"))
```

@@

```{r}
g_1ha_cbplot + g_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "growth_both_spei.png"))
```

@@

combine
```{r}
f_1ha_cbplot + f_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "flwr_both_spei.png"))
```

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, paged.print=FALSE)
library(tidyverse)
library(here)
library(mgcv)
library(dlnm)
library(bbmle)
library(lubridate)
library(glue)
library(patchwork)
# library(sf)
library(tsibble)
library(broom)
library(gratia)
library(latex2exp)

@@

@@

```{r}
demog_full <- read_rds(here("analysis", "data", "derived_data", "ha_spei_combined.rds"))
```

@@

@@

combine
```{r}
s_1ha_cbplot + s_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "surv_both_spei.png"), width = 10, height = 5)
```

@@

```{r}
g_1ha_cbplot + g_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "growth_both_spei.png"), width = 10, height = 5)
```

@@

combine
```{r}
f_1ha_cbplot + f_cf_cbplot + plot_layout(guides = "collect") + plot_annotation(tag_levels = "a", tag_suffix = ")")
ggsave(here("analysis", "figures", "flwr_both_spei.png"), width = 10, height = 5)
```
